---
title: "RWorksheet_calvario#4c.Rmd"
author: "Jolien"
date: "2024-11-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Use the dataset mpg

# a. Solutions on how to import a csv file into the environment.

```{r}
library(ggplot2)

mpg_data <- read.csv("mpg.csv")
str(mpg_data)
```

# b. Which variables from mpg dataset are categorical?

# The categorical variables in the mpg dataset are manufacturer, model, trans, drv, fl, and class. These represent distinct groups or categories, such as car brand, transmission type, drivetrain, fuel type, and car class.


# c. Which are continuous variables?

# The continuous variables in the mpg dataset are displ, cty, hwy, and cyl, as they represent measurable numerical values.


# 2.1. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.

# a. Group the manufacturers and find the unique models. Show your codes and result.

```{r}
library(dplyr)
manufacturer_model <- mpg %>%
  group_by(manufacturer) %>%
  summarize(model_num = n_distinct(model)) %>%
  arrange(desc(model_num))

manufacturer_model

variations_num <- table(mpg$model)
variations_num [variations_num == max(variations_num)]
```


# b. Graph the result by using plot() and ggplot(). Write the codes and its result.

```{r}
manufacturer_data <- setNames(
  manufacturer_model$model_num, 
  manufacturer_model$manufacturer
  )

barplot(manufacturer_data, 
        main = "Number of Models per Manufacturer",
        xlab = "Manufacturer",
        ylab = "Number of Models",
        col = c("violet", "pink", "lightgreen", "lavender", "lightblue"),
        las = 3)

variations_num <- mpg %>%
  group_by(model) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

variations_num

ggplot(variations_num, aes(x = reorder(model, count), y = count)) +
  geom_bar(stat = "identity", fill = "violet") + 
  coord_flip() +
  labs(title = "Number of Variations per Model", x = "Model", y = "Count") +
  theme_light()   
```

 
 

# 2.2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

# a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?

```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```

# ggplot(mpg, aes(model, manufacturer)) + geom_point() creates a scatter plot showing the relationship between car models (model) on the x-axis and manufacturers (manufacturer) on the y-axis, with each point representing a car observation.


# b. For you, is it useful? If not, how could you modify the data to make it more informative? 

# In its current form, this plot isn't very useful as it doesn't effectively visualize the relationship between car models and manufacturers. Since both variables are categorical, a scatter plot isn't the most appropriate way to represent this relationship.
 

# 3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.

```{r}
library(ggplot2)

# Top 20 observations from the mpg_data dataset
top_20_data <- head(mpg_data, 20)

# Plot using ggplot2
ggplot(top_20_data, aes(x = model, y = year)) +
  geom_point() +
  labs(
    title = "Model vs Year (Top 20 Observations)", 
    x = "Model", 
    y = "Year"
  ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(hjust = 0.5)  # Centering the title
  )
```


# 4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result

# a. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and colors. Show code and results.

```{r}
library(dplyr)
library(ggplot2)

car_counts <- mpg_data %>%
  group_by(model) %>%
  summarise(count = n())

top_20_models <- car_counts %>%
  arrange(desc(count)) %>%
  head(20)

ggplot(top_20_models, aes(x = reorder(model, -count), y = count)) +
  geom_bar(stat = "identity", fill = "violet") +
  labs(
    title = "Top 20 Models by Number of Cars",
    x = "Car Model",
    y = "Number of Cars"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels
```


# b. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.

```{r}
ggplot(top_20_models, aes(x = reorder(model, -count), y = count)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(
    title = "Top 20 Models by Number of Cars (Horizontal)",
    x = "Car Model",
    y = "Number of Cars"
  ) +
  coord_flip() +   
  theme_minimal()   
```

# 5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using geom_point with aesthetic color = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

# a. How would you describe its relationship? Show the codes and its result.

```{r}
library(ggplot2)

ggplot(mpg_data, aes(x = cyl, y = displ, color = displ)) +
  geom_point() + 
  labs(
    title = "Relationship between No. of Cylinders and Engine Displacement",
    x = "Number of Cylinders",
    y = "Engine Displacement (L)"
  ) +
  scale_color_gradient(low = "violet", high = "cyan") +    
  theme_minimal()
```

# The plot will show a scatter plot with the number of cylinders on the x-axis and the engine displacement on the y-axis. The points will be colored based on the engine displacement, where lower displacements will be shaded in violet, and higher displacements will be shaded in cyan.

